Fluent statistical computing interfaces for biological data analysis

### 2020-03-23 Rmd/05-embeddings.Rmd
adjusting posing queries para

We combine the tour view in a side by side layout with a scatter
plot view from an embedding algorithm. The views are linked; analysts can brush a group of points in the embedding view, and see its shape in the tour view.
This can be used for either manually identifying clusters, or verifying the location of clusters relative to other groupings. Consequently, simple linked
brushing can alleviate a downfall of non-linear methods such as UMAP or t-SNE
that break hierarchical structure of clusters. Aside from simple
linked brushing, we can also pose more complex queries via interrogating
the neighbourhood graph. Using a linked neighbourhood brush, we can 
visually investigate the nearest neighbour relationships in the 
high-dimensional space via brushing in the embedding view. The user can select
the number of nearest neighbours directly, and modify the distance metric 
used for determining the neighbours. This allows users to interrogate the stability of clusters generated in the embedding view. Multiple brushes
can be used to pose queries in either the tour view or the embedding view;
interface controls allow these brushes to combine using logical operators
such as 'and', 'or', or 'not'. The use of linked brushing goes beyond
simple color highlights, allowing analysts to get a more holistic
view of the effect of an embedding algorithm.

### Making comparisons

>>>>

We have combined the tour view in a side by side layout with 
a scatter plot view that represents the output of an embedding algorithm. 
These views are linked; analysts can brush regions or
highlight collections of points in either view. Linked highlighting can
be performed when points have been previously labeled according to some
discrete structure, i.e. cluster labels are available. This is achieved
via the analyst clicking on groups in the legend, which causes unclicked
groupings to have their points become less opaque. Consequently, 
simple linked highlighting can alleviate a known downfall of methods such as 
UMAP or t-SNE: that is distances between clusters are misleading. By
highlighting corresponding clusters in the tour view, the analyst can
see the relationship between clusters, and therefore obtain a more accurate 
representation of the topology of their data.

Simple linked brushing is achieved via mouse-click and drag movements. By
default, when brushing occurs in the tour view, the current projection is 
paused and corresponding points in the embedding view are highlighted. 
Likewise, when brushing occurs in the embedding view, corresponding points
in the tour view are highlighted. In this case, an analyst can use brushing
for manually identifying clusters and verifying cluster locations and shapes:
brushing in the embedding view gives analysts a sense of the shape and 
proximity of cluster in high-dimensional space.

### Making comparisons: think global, act local

Aside from simple
linked brushing, we can also pose more complex queries via interrogating
the neighbourhood graph. Using a linked neighbourhood brush, we can 
visually investigate the nearest neighbour relationships in the 
high-dimensional space via brushing in the embedding view. The user can select
the number of nearest neighbours directly, and modify the distance metric 
used for determining the neighbours. This allows users to interrogate the stability of clusters generated in the embedding view. Multiple brushes
can be used to pose queries in either the tour view or the embedding view;
interface controls allow these brushes to combine using logical operators
such as 'and', 'or', or 'not'. The use of linked brushing goes beyond
simple color highlights, allowing analysts to get a more holistic
view of the effect of an embedding algorithm.
